---
title: "Demographics Cleaning Practice"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Crystal Lewis
#### 11/12/2019


#### Install and call packages we need

Install packages
```{r, results='hide', warning=FALSE, message=FALSE, eval=FALSE}

install.packages(c("tidyverse", "lubridate","labelled", "dataMaid", "janitor"))
```

Call the packages
```{r, results='hide', warning=FALSE, message=FALSE}

library(tidyverse)
library(lubridate)
library(labelled)
```

What does our data currently look like?
```{r,echo=FALSE, message=FALSE}

Demo <- read_csv("https://raw.githubusercontent.com/Cghlewis/R-Ladies-STL-Cleaning-Data-R/master/Data/Demographics.csv")
knitr::kable(head(Demo, 3))
```

What do we want our data to look like?
```{r,echo=FALSE, message=FALSE}

Demofinal <- read_csv("https://raw.githubusercontent.com/Cghlewis/R-Ladies-STL-Cleaning-Data-R/master/Data/Clean_Demographics_2018_19.csv")
knitr::kable(head(Demofinal, 3))
```

##### My plan
###### (1) Read the data in
###### (2) Select variables to keep
###### (3) Rename variables
###### (4) Filter out Students with NA Student IDs
###### (5) Check for any duplicate rows
###### (6) Transform the School District column
###### (7) Remove the % sign from the Fall_Attendance column
###### (8) Recode Race and Gender
###### (9) Add variable labels

Read in our demographics file using `read_csv`
```{r, results='hide', warning=FALSE, message=FALSE}
Demo <- read_csv('https://raw.githubusercontent.com/Cghlewis/R-Ladies-STL-Cleaning-Data-R/master/Data/Demographics.csv', na="*")
```

Drop StudentName and ZipCode using `select`
```{r, warning=FALSE, message=FALSE}

names(Demo)
Demo <- Demo%>%select(-`Student Name`, -ZipCode)
```

Rename Columns using `setNames`
```{r, results='hide', warning=FALSE, message=FALSE}

names(Demo)
Demo <- Demo %>% setNames(c("School_District", "Student_ID","Race", "Gender",
                                "Fall_Attendance", "Spring_Attendance"))
```

Filter out any NA StudentID using `filter`
```{r, warning=FALSE, message=FALSE}

summary(Demo$Student_ID)

Demo <- Demo %>% filter(!is.na(Student_ID))

summary(Demo$Student_ID)
```

Check for duplicates and remove using `distinct`
```{r, warning=FALSE, message=FALSE}

dim(Demo)

distinct(Demo) %>% count()

Demo <- distinct(Demo)

dim(Demo)
```

Transform School_District using `separate` 
```{r, warning=FALSE, message=FALSE}

head(as.data.frame(Demo),2)
Demo <- Demo %>% separate(School_District,into=c("School","District"), sep='-')
head(as.data.frame(Demo),2)

```

Change Fall_Attendance class into numeric using `as.numeric` and remove the % using `str_remove`
```{r, warning=FALSE, message=FALSE}

head(as.data.frame(Demo),2)
class(Demo$Fall_Attendance)
Demo <- Demo %>% mutate(Fall_Attendance=as.numeric(str_remove(Fall_Attendance,"%")))
head(as.data.frame(Demo),2)
```
Last, recode variables.

Change Race class into a factor and recode values using `recode_factor`
```{r, warning=FALSE, message=FALSE}

class(Demo$Race)
table(Demo$Race)

Demo<-Demo%>%mutate(Race=recode_factor(Race, 'W'="White", 'I'="American Indian",
                                       'A'="Asian", 'B'="Black", 'H'="Latino/Hispanic",
                                       'P'="Native Hawaiian/Pacific Islander"))

levels(Demo$Race)
janitor::tabyl(Demo$Race)
```

Change Gender class into factor and recode values using `recode_factor`
```{r, warning=FALSE, message=FALSE}

class(Demo$Gender)
table(Demo$Gender)

Demo<-Demo%>%mutate(Gender=recode_factor(Gender,`1`="Male", `2`="Female"))

levels(Demo$Gender)
```

Add variable labels using `var_label`
```{r, warning=FALSE, message=FALSE}

names(Demo)

var_label(Demo) <- list(School = "SchoolName",District="DistrictName", 
                                Student_ID="Student ID", Race="Student Race", 
                                Gender="Student Gender", Fall_Attendance="Percent Attendance Fall", 
                                Spring_Attendance="Percent Attendance Spring")
```

##### While the code above works just fine, we could also write code chunks together using %>%. Then we only have to call the data once rather than 8 times.
```{r, results='hide', warning=FALSE, message=FALSE, eval=FALSE}

Demo <- Demo %>% select(-`Student Name`, -ZipCode)%>%
  setNames(c("School_District", "Student_ID","Race", "Gender",
            "Fall_Attendance", "Spring_Attendance")) %>%
  filter(!is.na(Student_ID)) %>% distinct() %>%
  separate(School_District,into=c("School","District"), sep='-') %>%
  mutate(Fall_Attendance=as.numeric(str_remove(Fall_Attendance,"%")),
                                    Race=recode_factor(Race, 'W'="White", 'I'="American Indian",
                                       'A'="Asian", 'B'="Black", 'H'="Latino/Hispanic",
                                       'P'="Native Hawaiian/Pacific Islander"),
                                    Gender=recode_factor(Gender,`1`="Male", `2`="Female"))
```

#### Last we can make a codebook of variable descriptives
Make a codebook using `makeCodebook`
```{r, results='hide', warning=FALSE, message=FALSE, eval=FALSE}

dataMaid::makeCodebook(Demo, reportTitle="Codebook for R-ladies Demographic Data")
```
#### And we can export our data to share with others or analyze in other programs.
Export Data using `write_csv`
```{r, results='hide', warning=FALSE, message=FALSE, eval=FALSE}

write_csv(Demo,"Clean_Demographics_2018_19.csv")
```
#### Also, if you wanted your data in tidy format, you could restructure from wide to long
```{r, warning=FALSE, message=FALSE}

DemoLong<-Demo%>%pivot_longer(cols=contains("Attend"), 
                                          names_to=c("Time",".value"), names_sep="_")

head(as.data.frame(DemoLong))
```
