---
title: "Discipline Cleaning Practice"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Crystal Lewis
#### 11/12/2019


#### Install and call packages we need

Install packages
```{r, results='hide', warning=FALSE, message=FALSE, eval=FALSE}
install.packages(c("tidyverse", "lubridate","labelled", "dataMaid"))
```

Call the packages
```{r, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(labelled)
```

What does our data currently look like?
```{r,echo=FALSE, message=FALSE}
ISS<-read_csv("https://raw.githubusercontent.com/Cghlewis/R-Ladies-STL-Cleaning-Data-R/master/Data/Discipline.csv")
knitr::kable(head(ISS, 3))
```

What do we want our data to look like?
```{r,echo=FALSE, message=FALSE}
ISSFinal<-read_csv("https://raw.githubusercontent.com/Cghlewis/R-Ladies-STL-Cleaning-Data-R/master/Data/Clean_Discipline_2018_19.csv")
knitr::kable(head(ISSFinal, 3))
```

##### My plan
###### (1) Read in the file
###### (2) Drop unncessary variables
###### (3) Rename columns
###### (4) Create a mont column
###### (5) Create an ISS column
###### (6) Filter to only ISS cases
###### (7) Group by StudentID and Month and count ISS incidents for Fall and Spring
###### (8) Merge the Fall and Spring together
###### (9) Fill NA ISS with zeros
###### (10) Add variable labels

Read in our discipline file using `read_csv`
```{r, results='hide', warning=FALSE, message=FALSE}
ISS<-read_csv('https://raw.githubusercontent.com/Cghlewis/R-Ladies-STL-Cleaning-Data-R/master/Data/Discipline.csv')
```
Drop Incident Description and Incident Place using `select`
```{r, warning=FALSE, message=FALSE}
names(ISS)
ISS<-ISS%>%select(-contains("Incident"))
```
Rename the Student ID column using `rename`
```{r, results='hide', warning=FALSE, message=FALSE}
names(ISS)
ISS<-ISS%>%rename(Student_ID=`Student ID`)
```
Create a month column using `mutate`, `mdy` and `month`
```{r, warning=FALSE, message=FALSE}
glimpse(ISS)
ISS<-ISS%>%mutate(Month=month(mdy(Date)))
head(as.data.frame(ISS))
```
Create a column with just ISS using `str_extract`
```{r, warning=FALSE, message=FALSE}
ISS<-ISS%>%mutate(ISSvar=str_extract(Action, "ISS"))
head(as.data.frame(ISS))
```
Filter to keep only the rows where ISS occurred using `filter`
```{r, warning=FALSE, message=FALSE}
ISS<-ISS%>%filter(ISSvar=="ISS")
```
Count the number of ISS per Student for fall/spring using `filter`, `group_by` and `summarise`
```{r, warning=FALSE, message=FALSE}
ISS_S<-ISS%>%filter(Month<6)%>%group_by(Student_ID)%>%
  summarise(Spring_ISS=n())

ISS_F<-ISS%>%filter(Month>7)%>%group_by(Student_ID)%>%
  summarise(Fall_ISS=n())

head(as.data.frame(ISS_S))
head(as.data.frame(ISS_F))
```
Merge the two data together
```{r, warning=FALSE, message=FALSE}
ISSJoin<-full_join(ISS_F, ISS_S, by="Student_ID")
head(as.data.frame(ISSJoin))
```
Fill NA ISS with 0s using `replace_na`
```{r, warning=FALSE, message=FALSE}
ISSJoin<-ISSJoin%>%tidyr::replace_na(list(Fall_ISS=0, Spring_ISS=0))
```
Add variable labels using `var_label`
```{r, warning=FALSE, message=FALSE}
names(ISSJoin)
var_label(ISSJoin) <- list(Student_ID="Student ID", Fall_ISS="Fall ISS referrals",
                        Spring_ISS="Spring ISS referrals")
```

##### While the code above works just fine, we could also write code chunks 2-6 together using %>%. Then we only have to call the data once rather than 5 times.
```{r, results='hide', warning=FALSE, message=FALSE, eval=FALSE}
ISS<-ISS%>%select(-contains("Incident"))%>%
  rename(Student_ID=`Student ID`)%>%
  mutate(Month=month(mdy(Date)), 
         ISSvar=str_extract(Action, "ISS"))%>%
  filter(ISSvar=="ISS")
```

#### Last we can make a codebook of variable descriptives
Make a codebook using `makeCodebook`
```{r, results='hide', warning=FALSE, message=FALSE, eval=FALSE}
dataMaid::makeCodebook(ISSJoin, reportTitle="Codebook for R-ladies Discipline Data")
```
#### And we can export our data to share with others or analyze in other programs.
Export Data using `write_csv`
```{r, results='hide', warning=FALSE, message=FALSE, eval=FALSE}
write_csv(ISSJoin,"Clean_Discipline_2018_19.csv")
```
#### Also, if you wanted your data in tidy format, you could restructure from wide to long
```{r, warning=FALSE, message=FALSE}
ISSLong<-ISSJoin%>%pivot_longer(cols=contains("ISS"), 
                              names_to=c("Time",".value"), names_sep="_")

head(as.data.frame(ISSLong))
```