---
title: "Cleaning Study Data and Merging Practice"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### Crystal Lewis
#### 11/12/2019


#### Install and call packages we need

Install packages
```{r, results='hide', warning=FALSE, message=FALSE, eval=FALSE}
install.packages(c("tidyverse", "lubridate","labelled", "dataMaid", "janitor"))
```

Call the packages
```{r, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(labelled)
```

Let's look at our treatment data
```{r,echo=FALSE, message=FALSE}
StudyData <- data.frame("Study_ID" = 1:3, 
                        "Student_ID"=c(71199755559,73951519331, 71397595973),
                        "Treatment"=c(0,1,1),
                        "LikertQ1" = c(1,4,3), "LikertQ2"=c(2,3,3), "LikertQ3"=c(1,1,4))
knitr::kable(head(StudyData))
summary(StudyData)
```

##### My plan to clean up the treatment data
###### (1) Recode Likert questions and add labels
###### (2) Add labels to treatment vars
###### (3) Add variable labels

Q1 just needs value labels using `labelled`
```{r, warning=FALSE, message=FALSE}
StudyData$LikertQ1<-labelled (StudyData$LikertQ1, c("Strongly Disagree"=1, "Disagree"=2,
                                                    "Agree"=3, "Strongly Agree"=4))
val_labels(StudyData$LikertQ1)
```
Q2 just needs value labels using `labelled`
```{r, warning=FALSE, message=FALSE}
StudyData$LikertQ2<-labelled (StudyData$LikertQ2, c("Strongly Disagree"=1, "Disagree"=2,
                                                    "Agree"=3, "Strongly Agree"=4))
val_labels(StudyData$LikertQ2)
```
Q3 needs to be reverse coded and labels added using `recode` and `labelled`
```{r, warning=FALSE, message=FALSE}
StudyData<-StudyData%>%mutate(LikertQ3.R=recode(LikertQ3, `1`=4, `2`=3, `3`=2, `4`=1))

StudyData$LikertQ3.R<-labelled (StudyData$LikertQ1, c("Strongly Disagree"=4, "Disagree"=3,
                                                    "Agree"=2, "Strongly Agree"=1))

table(StudyData$LikertQ3.R, StudyData$LikertQ3)

StudyData<-StudyData%>%select(-LikertQ3)
```
Add value labels to treatment using `labelled`
```{r, warning=FALSE, message=FALSE}
StudyData$Treatment<-labelled (StudyData$Treatment, c("Control"=0, "Treatment"=1))
```
Add variable labels using `labelled`
```{r, warning=FALSE, message=FALSE}
var_label(StudyData) <- list(Study_ID = "Study ID",Student_ID="District ID", 
                        LikertQ1="I like school?", LikertQ2="I like my teacher", 
                        LikertQ3.R="I don't want to come to school", Treatment="Treatment status")

```

##### Read in discipline and demographics

Read in our two district files that are cleaned
```{r, warning=FALSE, message=FALSE}
Demo<-read_csv("https://raw.githubusercontent.com/Cghlewis/R-Ladies-STL-Cleaning-Data-R/master/Data/Clean_Demographics_2018_19.csv")
Discipline<-read_csv("https://raw.githubusercontent.com/Cghlewis/R-Ladies-STL-Cleaning-Data-R/master/Data/Clean_Discipline_2018_19.csv")
```

Merge study data with district data
```{r, warning=FALSE, message=FALSE}
StudyDataComplete<-StudyData%>%left_join(Demo,by="Student_ID")%>%
  left_join(Discipline, by="Student_ID")

knitr::kable(head(StudyDataComplete))
```
